Spatial Reasoning and Statistical Graphics

### 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
formula massages
Fix scaled score boxplot

score.summary <- melt(ans.summary, measure.vars=20:24, value.name = "Score", variable.name = "Test")
score.summary$Test <- factor(score.summary$Test, levels=c("lineup", "card_rot", "fig_class", "folding", "vis_search"), labels=c("Lineups", "Card Rotation", "Figure Class.", "Paper Folding", "Visual Search"))
qrange$sqrt.k <- sqrt(qrange$k-1)
qrange$Test <- factor(qrange$testtype, levels=c("lineup", "card_rot", "fig_class", "folding", "vis_search"), labels=c("Lineups", "Card Rotation", "Figure Class.", "Paper Folding", "Visual Search"))

qr1 <- qrange
qr1$Type <- "Theoretical\nRange"
qr2 <- qrange
qr2$Type <- "Scores indicating\nguessing"
qr2$n <- with(qr2, -n/(k-1))

@@

ggplot() + 
  geom_bar(data=qrange, aes(width=.9, x=Test, y=n, fill=Type), alpha=.35, stat="identity", position="identity", colour="red", inherit.aes=F) +
  scale_fill_manual("", values=c(NA, "red")) + 
  geom_boxplot(data=score.summary, aes(x=factor(Test), y=Score), inherit.aes=F, fill="grey") +
  theme_bw()  + 
  ylab("Test Scores") + 
  xlab("") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
  theme(plot.margin=unit(c(0,0,0,0),"mm")) + 
  theme(legend.position=c(.80,.825),  legend.box.just = "left", 
        legend.key.height=unit(8, "mm")) 
  #coord_flip()
@
%

@@

This scaling gets slightly more complicated for scores which have two sub-groups, as with the figure classification test, which separately sumarizes male and female participants' scores. 
To get a single unified score with standard deviation, we completed the following calculations: 
\begin{align}
\mu_{\text{all}} &= (N_F\mu_{F} + N_M\mu_{M})/(N_F + N_M)\\
\sigma_{\text{all}} &= \sqrt{\left(N_F\sigma_F^2 + N_M\sigma_M^2\right)/(N_F + N_M)},
\end{align}
\newtext{where $\mu_F$ and $\mu_M$ are the mean of female and male scores respectively, $N_F$ and $N_M$ are the number of participants in each group, and $\sigma_F^2$ and $\sigma_M^2$ are the variance of each group.
Substituting in the provided numbers, }
\begin{align}
\mu_{\text{all}} &= \left(323\!*\!114.9\! +\! 294\!*\!120.0\right)/(323\!+\!294) \\
& = \Sexpr{round((120.0*294 + 114.9*323)/617/2, 1)}\\
\sigma_{\text{all}} &= \sqrt{\left(323(27.8)^2 \!+\! 294(30)^2\right)/(323\!+\!294)} \\
& = \Sexpr{round(sqrt((30^2*294 + 27.8^2*323)/617)/2, 1)}
\end{align}

@@

\begin{eqnarray}
\mu_{\text{part I}} &= 1/2 \mu_{\text{all}}\\
\sigma_{\text{part I}} &= 1/\sqrt{2} \sigma_{\text{all}}
\end{eqnarray}

@@

>>>>

score.summary <- melt(ans.summary, measure.vars=21:25, value.name = "Score", variable.name = "Test")
score.summary$Test <- factor(score.summary$Test, levels=c("lineup", "card_rot", "fig_class", "folding", "vis_search"), labels=c("Lineups", "Card Rotation", "Figure Class.", "Paper Folding", "Visual Search"))
qrange$sqrt.k <- sqrt(qrange$k-1)
qrange$Test <- factor(qrange$testtype, levels=c("lineup", "card_rot", "fig_class", "folding", "vis_search"), labels=c("Lineups", "Card Rotation", "Figure Class.", "Paper Folding", "Visual Search"))

qr1 <- qrange
qr1$Type <- "Theoretical Range"
qr2 <- qrange
qr2$Type <- "Scores indicating guessing"
qr2$n <- with(qr2, -n/(k-1))

@@

ggplot() + 
  geom_bar(data=qrange, aes(width=.9, x=Test, y=n, fill=Type), alpha=.35, stat="identity", position="identity", colour="red", inherit.aes=F) +
  scale_fill_manual("", values=c(NA, "red")) + 
  geom_boxplot(data=score.summary, aes(x=factor(Test), y=Score), inherit.aes=F, fill="grey") +
  theme_bw()  + 
  ylab("Test Scores") + 
  xlab("") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
  theme(plot.margin=unit(c(0,0,0,0),"mm")) + 
  theme(legend.position=c(.80,.125),  legend.box.just = "left", 
        legend.key.height=unit(8, "mm")) 
  #coord_flip()
@
%

@@

This scaling gets slightly more complicated for scores which have two sub-groups, as with the figure classification test, which separately sumarizes male and female participants' scores. 
To get a single unified score with standard deviation, we completed the following calculations: 
\begin{align}
\mu_{\text{all}} &= (N_F\mu_{F} + N_M\mu_{M})/(N_F + N_M)\\
\sigma_{\text{all}} &= \sqrt{\left(N_F\sigma_F^2 + N_M\sigma_M^2\right)/(N_F + N_M)},
\end{align}
\newtext{where $\mu_F$ and $\mu_M$ are the mean of female and male scores respectively, $N_F$ and $N_M$ are the number of participants in each group, and $\sigma_F^2$ and $\sigma_M^2$ are the variance of each group.
Substituting in the provided numbers, we get}
\begin{align*}
\mu_{\text{all}} &= \left(323\!*\!114.9\! +\! 294\!*\!120.0\right)/(323\!+\!294) \\
& = \Sexpr{round((120.0*294 + 114.9*323)/617/2, 1)}\\
\sigma_{\text{all}} &= \sqrt{\left(323(27.8)^2 \!+\! 294(30)^2\right)/(323\!+\!294)} \\
& = \Sexpr{round(sqrt((30^2*294 + 27.8^2*323)/617)/2, 1)}.
\end{align*}

@@

\begin{eqnarray*}
\mu_{\text{part I}} &= 1/2 \cdot \mu_{\text{all}}\\
\sigma_{\text{part I}} &= 1/\sqrt{2} \cdot \sigma_{\text{all}}
\end{eqnarray*}

@@

xtabs(~stem+sex, data=ans.model)
# sad day!
